---
title: "Data_exploration"
author: "Allen Bush"
date: "2023-05-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls()) # clear workspace

library(tidyverse)
library(tidylog)
library(readxl)
library(ggtext)
library(MetBrewer)


pest_conc <- read_excel("master_pesticides.xlsx")
pest_detect <- read_excel("pesticide_detection.xlsx")


```
Code inspired (practically copy-pasted) by Andrew Heiss at: https://www.andrewheiss.com/blog/2022/05/09/hurdle-lognormal-gaussian-brms/#hurdle-lognormal-model

```{r}
# Use the Johnson color palette
clrs <- MetBrewer::met.brewer("Johnson")


# Custom ggplot theme to make pretty plots
# Get the font at https://fonts.google.com/specimen/Jost
theme_nice <- function() {
  theme_minimal(base_family = "Jost") +
    theme(panel.grid.minor = element_blank(),
          plot.title = element_text(family = "Jost", face = "bold"),
          axis.title = element_text(family = "Jost Medium"),
          strip.text = element_text(family = "Jost", face = "bold",
                                    size = rel(1), hjust = 0),
          strip.background = element_rect(fill = "grey80", color = NA))
}
```


Data manipulation
Set up pest_conc concentration in long format rather than having one column per pesticide
convert nd to zeros, <LOQ to LOD as brms prefers the lower bound of the censored data
```{r}
pest_conc_long <- pest_conc %>% pivot_longer(cols = one.naphthol:fludioxonil, names_to = "pesticide", values_to = "recorded_concentration")

pest_master_long <- right_join(pest_conc_long, pest_detect, by = "pesticide") # tidylog says there are duplicates - identify these

sum(duplicated(pest_master_long)) # 0 so tidylog must not be sending the message I thought it was

pest_master_long <- pest_master_long %>%
  mutate(concentration = as.numeric(ifelse(recorded_concentration == "nd", 0, 
                                            ifelse(recorded_concentration == "<LOQ", LOD, recorded_concentration))), # NAs introduced by coercion -> from samples in 2013 where recorded_concentration == NA just remove for now
         status = ifelse(recorded_concentration == "nd", "nd", 
                                            ifelse(recorded_concentration == "<LOQ", "<LOQ", "quantified"))) %>%
  drop_na(concentration) %>%
  mutate(is_zero = concentration == 0)

pest_conc_quant <- subset(pest_master_long, status == "quantified")
```


Proportion of zeros
```{r}
pest_master_long %>%
  #group_by(pesticide) %>%
  count(is_zero) %>%
  mutate(prop = n/sum(n))
```



Data distribution for quantified pesticides
```{r}
pest_conc_quant %>%
  ggplot() +
  geom_histogram(aes(x = concentration), binwidth = 0.1) +
  facet_wrap(~pesticide, scales = "free") +
  theme_classic()
```

