---
title: "simulating partial pooling and censoring"
author: "Allen Bush"
date: "2023-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls()) # clear workspace

library(tidyverse)
library(brms)
library(cmdstanr)
library(tidybayes)
library(ggridges)
library(tidylog)
library(readxl)
library(broom)
library(posterior)
library(ggdist)

pest_conc <- read_excel("master_pesticides.xlsx")
pest_detect <- read_excel("pesticide_detection.xlsx")


```

Data manipulation
```{r}
pest_conc_long <- pest_conc %>% pivot_longer(cols = one.naphthol:fludioxonil, names_to = "pesticide", values_to = "recorded_concentration")

pest_master_long <- right_join(pest_conc_long, pest_detect, by = "pesticide")

pest_master_long <- pest_master_long %>%
  mutate(concentration = as.numeric(ifelse(recorded_concentration == "nd", 0, 
                                            ifelse(recorded_concentration == "<LOQ", LOD, recorded_concentration))), # NAs introduced by coercion -> from samples in 2013 where recorded_concentration == NA just remove for now
         status = ifelse(recorded_concentration == "nd", "nd", 
                                            ifelse(recorded_concentration == "<LOQ", "<LOQ", "quantified"))) %>%
  drop_na(concentration)

pest_conc_quant <- subset(pest_master_long, status == "quantified")
```
# Part 1. recover parameters from hurdle lognormal for one pesticide, 5 parameter values, different sample sizes

##simulate censored hurdle lognormal for generic pesticide with n = 2000
```{r}
set.seed(333)

# sample size
n1 <- 2000 

# probability of a zero
pi <- 0.65
# lognormal mean & sigma
mu_log <- log(7)
sigma_log <- log(2)

# simulate data
y <- (1 - rbinom(n1, 1, prob = pi)) * rlnorm(n1, mu_log, sigma_log)

ggplot(data = tibble(y)) +
  geom_histogram(aes(y))

```

###adapted from: https://bookdown.org/content/3686/tools-in-the-trunk.html#censored-data-in-jags-brms
```{r}
d <- tibble(y = y)


# values below 3 are censored
t1 <- 3


d <-
  d %>% 
  mutate(y1   = if_else(y < t1, t1, y),
         cen1 = if_else(y < t1, "left", "none")) %>%
   mutate(y_na = ifelse(cen1 == "none", y, NA))

d
```

Histograms
```{r}
d %>% 
  pivot_longer(-cen1) %>% 
  mutate(name = factor(name, levels = c("y", "y1", "y_na"))) %>% 
  
  ggplot(aes(x = value)) +
  geom_histogram(linewidth = 0.25, binwidth = 1, 
                 fill = "skyblue4", color = "white") +
  xlab(NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  facet_wrap(~ name, ncol = 3)
```


Interval censoring from 3 to 5
```{r}
t2 <-3

t3 <- 5


d <-
  d %>% 
  mutate(y2   = if_else(y >= t2 & y < t3, t2, y),
         y3   = if_else(y >= t2 & y < t3, t3, y),
         cen2 = if_else(y >= t2 & y < t3, "interval", "none")) %>% 
  mutate(y_na = ifelse(cen2 == "none", y, NA))%>% 
  mutate(y4   = if_else(y >= t2 & y < t3, t2, 
                        if_else(y < t1, t1, y)),
         y5   = if_else(y >= t2 & y < t3, t3, y),
         cen3 = if_else(y >= t2 & y < t3, "interval",
                       if_else(y < t1, "left", "none"))) %>% 
  mutate(y_na = ifelse(cen3 == "none", y, NA))

d
```


Histograms
```{r}
d %>% 
  pivot_longer(cols = c(y, y4, y5, y_na)) %>% 
  mutate(name = factor(name, levels = c("y", "y4", "y5", "y_na"))) %>% 
  
  ggplot(aes(x = value)) +
  geom_histogram(linewidth = 0.25, binwidth = 0.5, 
                 fill = "skyblue4", color = "white") +
  labs(subtitle = "Our data have been updated, again",
       x = NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  facet_wrap(~ name, ncol = 2)
```


Model adapted from both previous links - set priors
```{r}
cens_hurdle_bf <- bf(y4 | cens(cen3, y5) ~ 1,
                     hu ~ 1,
                     family = "hurdle_lognormal"
                     )

get_prior(cens_hurdle_bf, data = d)

#logistic distribution for hu
curve(plogis(x), from = -5, to = 5)

cens_hurdle_priors <- c(
  prior(normal(1,0.5), class = "Intercept", dpar = "hu"),
  prior(normal(3,1.5), class = "Intercept"),
  prior(exponential(1), class = "sigma")
)

prior_predictions <-
  brm(data = d,
      formula = cens_hurdle_bf,
      prior = cens_hurdle_priors,
      sample_prior = "only",
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 12,
      backend = "cmdstanr")
print(prior_predictions)

d %>% 
  add_predicted_draws(prior_predictions, ndraws = 6) |> 
  ggplot(aes(x = y4, y = .prediction)) + geom_point() + facet_wrap(~.draw, scales = "free")
```


Recover parameters
```{r}
fit1 <-  brm(data = d,
      formula = cens_hurdle_bf,
      prior = cens_hurdle_priors,
      sample_prior = "yes",
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 12,
      backend = "cmdstanr")

print(fit1)

draws<- d %>%
  add_predicted_draws(fit1, ndraws = 50)


hurdle.actualdata <- ggplot(data = draws) +
  geom_histogram(aes(x = y, y = ..density..),  
                 fill = "skyblue4", color = "white", binwidth = 1) +
  geom_density(aes( x = .prediction,  group = .draw),
            size  = 1/4, alpha = 1/3, color = "lightblue") +
  labs(title = "Actual Data") +
  theme_classic()
hurdle.actualdata

hurdle.censoreddata <- ggplot(data = draws) +
  geom_histogram(aes(x = y4, y = ..density..),  
                 fill = "skyblue4", color = "white", binwidth = 1) +
  geom_density(aes( x = .prediction,  group = .draw),
            size  = 1/4, alpha = 1/3, color = "lightblue") +
  labs(title = "Censored Data") +
  theme_classic()

hurdle.censoreddata 


#spread_rvars to recover actual data
```
n = 2000 works very well to recover parameters
2000*probzero(0.65) = 1300 - > 700 actually "measured" points which is quite a lot compared to the actual data

I think we should keep n = 2000 but vary hu between 0.85 to 0.99 (by .1)

First, let's try with values from a real pesticide - Atrazine


```{r}
set.seed(333)

# sample size
n1 <- 2000 
atrazine_detect <- pest_detect %>%
  subset(pesticide == "atrazine")
hu_atrazine <- 0.9
dist_from_LOQ_atrazine <- 1.1
mu_atrazine <- atrazine_detect$LOQ * dist_from_LOQ_atrazine
sigma_atrazine <- 2
left_censor_atrazine <- atrazine_detect$LOD
interval_censor_atrazine <- atrazine_detect$LOQ

# simulate data
# need to convert every value under LOD to LOD,
# every value between LOD and LOQ to LOD (the lower bound)
# need a variable y_upper to specify the upper bound of the interval censoring (LOQ) 
atrazine_sim <- tibble(
  y = (1 - rbinom(n1, 1, prob = hu_atrazine)) * rlnorm(n1, mu_atrazine, sigma_atrazine)
) %>%
  mutate(y_cen = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, left_censor_atrazine, # every value between LOD and LOQ to LOD (the lower bound)
                        ifelse(y < left_censor_atrazine, left_censor_atrazine, y)), # every value under LOD to LOD
         
         upper_int_cens = ifelse( y >= left_censor_atrazine & y < interval_censor_atrazine, interval_censor_atrazine, y)) %>% #specify upper bound of interval censoring
  
  mutate(censoring = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, "interval", 
                            ifelse(y < left_censor_atrazine, "left", "none"))) %>% # specify type of censoring
  mutate(y_na = ifelse(censoring == "none", y, NA))



atrazine_sim %>% 
  pivot_longer(cols = c(y, y_cen, upper_int_cens, y_na)) %>% 
  mutate(name = factor(name, levels = c("y", "y_cen", "upper_int_cens", "y_na"))) %>% 
  
  ggplot(aes(x = value)) +
  geom_histogram(linewidth = 0.25, binwidth = 0.1, 
                 fill = "skyblue4", color = "white") +
  labs(x = NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  facet_wrap(~ name, ncol = 2) + 
  coord_cartesian(xlim = c(0, 10), ylim = c(0,100))

tibble(
y = rlnorm(n1, mu_atrazine, sigma_atrazine)) %>%
    ggplot(aes(x = y)) +
  geom_histogram(linewidth = 0.25, binwidth = 0.1, 
                 fill = "skyblue4") 

mean(rlnorm(n1, mu_atrazine, sigma_atrazine))
```



```{r}
atrazine_sim2 <- tibble(
  y = (1 - rbinom(n1, 1, prob = hu_atrazine)) * rlnorm(n1, calculate_a_lnorm(mu_atrazine, sigma_atrazine), calculate_b_lnorm(mu_atrazine, sigma_atrazine))
) %>%
  mutate(y_cen = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, left_censor_atrazine, # every value between LOD and LOQ to LOD (the lower bound)
                        ifelse(y < left_censor_atrazine, left_censor_atrazine, y)), # every value under LOD to LOD
         
         upper_int_cens = ifelse( y >= left_censor_atrazine & y < interval_censor_atrazine, interval_censor_atrazine, y)) %>% #specify upper bound of interval censoring
  
  mutate(censoring = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, "interval", 
                            ifelse(y < left_censor_atrazine, "left", "none"))) %>% # specify type of censoring
  mutate(y_na = ifelse(censoring == "none", y, NA))


atrazine_sim2 %>% 
  pivot_longer(cols = c(y, y_cen, upper_int_cens, y_na)) %>% 
  mutate(name = factor(name, levels = c("y", "y_cen", "upper_int_cens", "y_na"))) %>% 
  
  ggplot(aes(x = value)) +
  geom_histogram(linewidth = 0.25, binwidth = 0.1, 
                 fill = "skyblue4", color = "white") +
  labs(x = NULL) +
  scale_y_continuous(NULL, breaks = NULL) +
  facet_wrap(~ name, ncol = 2) + 
  coord_cartesian(xlim = c(0, 10), ylim = c(0,100))

```


Atrazine model
Priors
```{r}
cens_hurdle_atrazine <- bf(y_cen | cens(censoring, upper_int_cens) ~ 1,
                           hu ~ 1,
                           family = "hurdle_lognormal"
)

get_prior(cens_hurdle_atrazine, data = atrazine_sim)

cens_hurdle_priors <- c(
  prior(normal(1,0.5), class = "Intercept", dpar = "hu"),
  prior(normal(3,1.5), class = "Intercept"),
  prior(exponential(1), class = "sigma")
)

mod_atrazine_priors <- brm(data = atrazine_sim,
                      formula = cens_hurdle_atrazine,
                      family = "hurdle_lognormal",
                      prior = cens_hurdle_priors,
                      sample_prior = "only",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

prior_draws <- as_draws_df(mod_atrazine_priors) %>% select(b_Intercept, b_hu_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept, sigma))

true_values <- tibble(
  name = as.factor(c("b_Intercept", "b_hu_Intercept", "sigma")),
  value = c(log(mu_atrazine), qlogis(hu_atrazine), log(sigma_atrazine))
)


ggplot() +
  stat_histinterval(data = prior_draws, aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
# need to make the priors wider for intercept and hu - I know, this is technically cheating
```

Priors pt.2
```{r}
cens_hurdle_priors2 <- c(
  prior(normal(1.5,1), class = "Intercept", dpar = "hu"),
  prior(normal(3,3), class = "Intercept"),
  prior(exponential(1), class = "sigma")
)

mod_atrazine_priors2 <- brm(data = atrazine_sim,
                      formula = cens_hurdle_atrazine,
                      family = "hurdle_lognormal",
                      prior = cens_hurdle_priors2,
                      sample_prior = "only",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

prior_draws2 <- as_draws_df(mod_atrazine_priors2) %>% select(b_Intercept, b_hu_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept, sigma))


ggplot() +
  stat_histinterval(data = prior_draws2, aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")

# okay, this is much better - will have to revise these later
```






```{r}


mod_atrazine_1 <- brm(data = atrazine_sim,
                      formula = cens_hurdle_atrazine,
                      prior = cens_hurdle_priors2,
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")
broom.mixed::tidy(mod_atrazine_1)

plot(mod_atrazine_1)


```

posterior predictions
```{r}
atrazine_draws <- atrazine_sim %>%
  add_predicted_draws(mod_atrazine_1, ndraws = 100) 



ggplot(atrazine_draws) +
  geom_point(aes(x = .prediction, y = y)) +
  geom_abline()



posterior_draws <- as_draws_df(mod_atrazine_1) %>% select(b_Intercept, b_hu_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept, sigma))


ggplot() +
  stat_histinterval(data = posterior_draws, aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```

Let's slow down and model the data without censoring

```{r}
uncens_hurdle_atrazine <- bf(y ~ 1,
                           hu ~ 1,
                           family = "hurdle_lognormal"
)

get_prior(uncens_hurdle_atrazine, data = atrazine_sim)

mod_atrazine_uncensored <- brm(data = atrazine_sim,
                      formula = uncens_hurdle_atrazine,
                      family = "hurdle_lognormal",
                      prior = cens_hurdle_priors2,
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")
broom.mixed::tidy(mod_atrazine_uncensored)

plot(mod_atrazine_uncensored)
```

posterior predictions
```{r}
atrazine_draws_uncensored <- atrazine_sim %>%
  add_predicted_draws(mod_atrazine_uncensored, ndraws = 100) 



ggplot(atrazine_draws_uncensored) +
  geom_point(aes(x = .prediction, y = y)) +
  geom_abline()



posterior_draws_uncensored <- as_draws_df(mod_atrazine_uncensored) %>% select(b_Intercept, b_hu_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept, sigma))


ggplot() +
  stat_histinterval(data = posterior_draws_uncensored, aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```


According to this topic on the stan forum https://discourse.mc-stan.org/t/link-sigma-for-lognormal-family/11026 , the link function only applies to sigma when there are variables used to predict sigma. Otherwise, sigma is on the response scale no matter the link function (?)
Also, I think that the previous model uses the response scale for the posterior mean of the lognormal. Let's try setting the link to log

```{r}
uncens_hurdle_atrazine2 <- bf(y ~ 1,
                           hu ~ 1,
                           family = hurdle_lognormal(link = "log", link_hu = "logit", link_sigma = "identity")
)


```
 weird - when data is uncensored, the link needs to be identity
 Let's see when the data is censored
 
```{r}
cens_hurdle_atrazine2 <- bf(y_cen | cens(censoring, upper_int_cens) ~ 1,
                           hu ~ 1,
                           family = hurdle_lognormal(link = "log", link_sigma = "identity", link_hu = "logit")
)
```
Still doesn't work 


```{r}
cens_hurdle_atrazine3 <- bf(y_cen | cens(censoring, upper_int_cens) ~ 1,
                           hu ~ 1,
                           family = hurdle_lognormal(link = "identity", link_sigma = "identity", link_hu = "logit")
)


get_prior(formula = cens_hurdle_atrazine3, data = atrazine_sim)

mod_atrazine_3 <- brm(data = atrazine_sim,
                      formula = cens_hurdle_atrazine,
                      prior = cens_hurdle_priors2,
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")
broom.mixed::tidy(mod_atrazine_3)

plot(mod_atrazine_3)
```

