---
title: "Understanding the lognormal"
author: "Allen Bush"
date: "2023-06-14"
output: html_document
---


Here, I will explore the parameterization of the lognormal distribution as it is confusing to me
I will simulate data from a lognormal distribution using the moment matching that Andrew MacDonald has shown me


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls()) # clear workspace

library(tidyverse)
library(brms)
library(cmdstanr)
library(tidybayes)
library(ggridges)
library(tidylog)
library(readxl)
library(broom)
library(posterior)
library(ggdist)
library(patchwork)


pest_detect <- read_excel("pesticide_detection.xlsx")

```

moment matching functions
https://discourse.mc-stan.org/t/lognormal-regression-and-moment-matching/26557 

a
 is the median of the lognormal distribution. Its also the mean of the normal distribution you’d get if you took the log of these lognormal values. 
b
 is the standard deviation of those values.
```{r}
calculate_a_lnorm <- function(mean, sd){

  log(mean^2/sqrt(sd^2 + mean^2))

}


calculate_b_lnorm <- function(mean, sd){

  sqrt(log(sd^2 / mean^2 + 1))

}
```

#Explore the lognormal with brms
```{r}
n <- 2000
mu <- 7
sd <- 1

a <- calculate_a_lnorm(mu, sd)
b <- calculate_b_lnorm(mu, sd)

```


Let's take a look at the Gaussian distribution with these parameters and compare with the lognormal
```{r}
data <- tibble(
  y = rnorm(n, mu, sd),
  log_y = log(y),
  log_norm = rlnorm(n, a, b),
  exp_log_norm = exp(log_norm)
)

y_hist <- ggplot(data) +
  geom_histogram(aes(x = y)) +
  geom_vline(xintercept = mean(data$y)) +
  labs(title = "y")

log_y_hist <- ggplot(data) +
  geom_histogram(aes(x = log_y)) +
  geom_vline(xintercept = mean(data$log_y)) +
  labs(title = "log y")

log_norm_hist <- ggplot(data) +
  geom_histogram(aes(x = log_norm)) +
  geom_vline(xintercept = mean(data$log_norm)) +
  labs(title = "log normal")

exp_log_norm_hist<- ggplot(data) +
  geom_histogram(aes(x = exp_log_norm)) +
  geom_vline(xintercept = mean(data$exp_log_norm)) +
  labs(title = "e ^ log normal")


y_hist + log_y_hist + log_norm_hist + exp_log_norm_hist

```

Okay, so the lognormal IS the normal with this moment matching 

Recover parameters
Gaussian model
```{r}
gaussian_model_bf <- bf(y ~ 1,
                        family = gaussian)


get_prior(gaussian_model_bf, data = data)

gaussian_priors <- c(
  prior(normal(7,1), class = "Intercept"),
  prior(exponential(1), class = "sigma")
)


gaussian_model <- brm(data = data,
                      formula = gaussian_model_bf,
                      prior = gaussian_priors,
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")
plot(gaussian_model)

broom.mixed::tidy(gaussian_model)

```

lognormal model
```{r}
lognormal_model_bf <- bf(log_norm ~ 1,
                        family = lognormal)


get_prior(lognormal_model_bf, data = data)

#let's go with the default priors for now as I am not sure if the priors are for a and b or mu and sd

lognormal_model <- brm(data = data,
                      formula = lognormal_model_bf,
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")
plot(lognormal_model)

broom.mixed::tidy(lognormal_model)

```


Sooooo... brm models parameters a and b

compare model predictions with lognormal data and normal data

```{r}
predicted_draws <- data %>%
  add_predicted_draws(lognormal_model, ndraws = 100) 



ppc_lognormal <- ggplot(predicted_draws) +
  geom_point(aes(x = .prediction, y = log_norm)) +
  geom_abline()

ppc_normal <- ggplot(predicted_draws) +
  geom_point(aes(x = .prediction, y = y)) +
  geom_abline()

ppc_lognormal + ppc_normal

pp_check(lognormal_model)
```

The model predictions are on the response scale (the mean mu of the normal distribution)


Add predictors
```{r}
paysage <- seq(from = -3, to = 3, length.out = n)

mu_pred <- mu + 0.5*paysage

a_pred <- calculate_a_lnorm(mean = mu_pred, sd = sd)

b_pred <- calculate_b_lnorm(mean = mu_pred, sd = sd)


data_pred <- tibble(
  paysage = seq(from = -3, to = 3, length.out = n),
  log_norm = rlnorm(n, a_pred, b_pred)
)

ggplot(data_pred, aes(x = paysage, y = log_norm)) +
  geom_point() +
  geom_smooth(method = "lm")
```

Recover parameters

```{r}
lognormal_pred_bf <- bf(log_norm ~ paysage,
                        family = lognormal)


get_prior(lognormal_pred_bf, data = data_pred)

#let's go with the default priors for now 

lognormal_pred_model <- brm(data = data_pred,
                      formula = lognormal_pred_bf,
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")
plot(lognormal_pred_model)
pp_check(lognormal_pred_model)
broom.mixed::tidy(lognormal_pred_model) 
```

predictions

```{r}
pred_mean <- data_pred %>%
  select(paysage) %>%
  add_epred_draws(lognormal_pred_model, ndraws = 100)


epred <- ggplot() + stat_lineribbon(data = pred_mean, aes(x = paysage, y = .epred)) + 
  geom_point(data = data_pred, aes(x = paysage, y = log_norm), pch = 21, fill = "orange", alpha = 0.2) + 
  scale_fill_brewer(palette = "Greens", direction = 1)


predictions <- data_pred %>%
  select(paysage) %>%
  add_predicted_draws(lognormal_pred_model, ndraws = 100)

predictions_plot <- ggplot() + stat_lineribbon(data = predictions, aes(x = paysage, y = .prediction)) + 
  geom_point(data = data_pred, aes(x = paysage, y = log_norm), pch = 21, fill = "orange", alpha = 0.2) + 
  scale_fill_brewer(palette = "Greens", direction = 1)

epred + predictions_plot
```

Okay, so I now understand how to add predictors and the predictions from the model seem to match the simulation
Now, to understand the value of 0.07 for betapaysage

```{r}
exp(0.07155) #1.074

```

Maybe a better way to model predictors would be to compute a and b first and then add the effect of the predictors

```{r}


data_pred2 <- tibble(
  paysage = seq(from = -3, to = 3, length.out = n),
  log_norm = rlnorm(n, a + 0.5*paysage, b_pred)
)

p1 <- ggplot(data_pred, aes(x = paysage, y = log_norm)) +
  geom_point() +
  geom_smooth(method = "lm")

p2 <-ggplot(data_pred2, aes(x = paysage, y = log_norm)) +
  geom_point() +
  geom_smooth(method = "lm")

p1 + p2
```

Adding the predictor on the lognormal scale is actually exponential
Let's see if 0.07 on the lognormal scale gives the same results as 0.5 on the response scale

```{r}
data_pred3 <- tibble(
  paysage = seq(from = -3, to = 3, length.out = n),
  log_norm = rlnorm(n, a + 0.07*paysage, b_pred)
)



p3 <-ggplot(data_pred3, aes(x = paysage, y = log_norm)) +
  geom_point() +
  geom_smooth(method = "lm")

p1 + p3
```

Approximately the same thing
So, adding a predictor to a rather than mu represents a percentage change on the response scale
To keep my life simple as the parameterization of the lognormal is extremely weird, I will add predictors on the log scale from now on


Let's see if we can recover the parameters for a random effect

```{r}
n_per_pesticide <- 200
pesticides <- c(1:20)
pesticide_deviation <- rnorm(n = length(pesticides), mean = 0, sd = log(1.2)) # pesticides vary by 40% below and above the mean (2*sd)

random_data <- tibble(
  pesticide_ID = pesticides,
  pesticide_effect = pesticide_deviation[pesticides]
) %>% 
  expand_grid(obs = 1:n_per_pesticide) %>%
  mutate(log_norm = rlnorm(length(obs), a + pesticide_effect, b_pred))
```

Model
```{r}
lognormal_random_bf <- bf(log_norm ~ (1|pesticide_ID),
                        family = lognormal)


get_prior(lognormal_random_bf, data = random_data)

#let's go with the default priors for now 

lognormal_random_model <- brm(data = random_data,
                      formula = lognormal_random_bf,
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")
plot(lognormal_random_model)
pp_check(lognormal_random_model)
print(lognormal_random_model) 
```

Plot pesticides
```{r}
pred_random <- random_data %>%
  select(pesticide_ID) %>%
  add_epred_draws(lognormal_random_model, ndraws = 100) %>%
  ungroup() %>% # by default, dataset is grouped by .row so won't be selected out in the next line of code
  select(pesticide_ID, .epred)

pred_random_mean<- random_data %>%
  select(log_norm) %>%
  add_epred_draws(lognormal_random_model, ndraws = 100, re_formula = NA) %>% # ignore random effect
  ungroup() %>%
  mutate(pesticide_ID = "mean") %>%
  select(pesticide_ID, .epred)

pred_random_full <- rbind(pred_random, pred_random_mean) %>%
  mutate(pesticide_ID = as.factor(pesticide_ID)) %>%
  mutate(pesticide_ID = fct_reorder(.f = pesticide_ID, .x = .epred))
  
ggplot(data = pred_random_full) +
  stat_slabinterval(aes(y = .epred, x = pesticide_ID, color = pesticide_ID)) +
  theme_classic() +
  theme(legend.position = "none")
```
7*1.4 = 9.8
7/1.4 = 5

Groups vary by 40% around the overall mean

So we now understand how to model random and fixed effects for the lognormal

Now, let's understand how to choose priors

The priors are for the a and b parameters

means should vary on the response scale between zero and nearly 1000
```{r}
priorsim <- tibble(
  mean = rnorm(5000, mean = 500, sd = 250),
  sdev = rexp(5000, 0.5),
  a = calculate_a_lnorm(mean, sdev),
  b = calculate_b_lnorm(mean, sdev)
) %>% pivot_longer(
  everything()
)


ggplot(data = priorsim, aes(x = value)) +
  geom_histogram() +
  facet_wrap(~name, scales = "free")
```

So, some wide priors for a and b could be:
a ~ N(6,2)
b ~ Exponential(2) aka mean = 1/2

model with priors

going to update lognormal_model;
lognormal_model_bf <- bf(log_norm ~ 1,
                        family = lognormal)
                        
lognormal_model <- brm(data = data,
                      formula = lognormal_model_bf,
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")
```{r}
get_prior(lognormal_model_bf, data = data)

lognormal_priors <- c(
  prior(normal(6,2), class = "Intercept"),
  prior(exponential(2), class = "sigma")
)


lognormal_model_w_priors <- update(lognormal_model, prior = lognormal_priors) # requires recompiling as the priors have changed but if we did not change the priors, the model would not need to recompile

plot(lognormal_model_w_priors)

broom.mixed::tidy(lognormal_model_w_priors)

```

The model output doesn't really change which is good news as we don't want the priors to have much of an influence

Okay, so the lognormal is definitely weird but we have a method to recover the proper parameters

Now, let's get funky by modeling a hurdle lognormal

The goal of the analysis is to fit a hurdle lognormal model to pesticide concentration data that are left censored under the limit of detection (LOD) and interval censored between LOD and the limit of quantification (LOQ).

We will then explore the effect of varying the proportion of zeros (hu) and the distance between the mean of the distribution and LOQ on the ability of the model to recover the parameters. I hypothesize that the closer the mean is to LOQ, the less precise the model should be as it has practically no observed data on the left side of the mean to infer the censored part of the distribution

We will use LOD and LOQ values for one pesticide (atrazine) and simulate a mean 40% further from LOQ for now

This will be the workflow to complexify the model:
1) simulate hurdle lognormal uncensored data and recover parameters
2) simulate lognormal left censored data and recover parameters
3) simulate lognormal interval censored data and recover parameters
4) simulate lognormal left + interval censored data and recover parameters
5) simulate hurdle lognormal left censored data and recover parameters
6) simulate hurdle lognormal interval censored data and recover parameters
7) bring it all together and simulate hurdle lognormal left and interval censored data and recover parameters

#1) simulate hurdle lognormal uncensored data and recover parameters

Simulate data
```{r}
set.seed(333)

n1 <- 2000 # sample size

atrazine_detect <- pest_detect %>%
  subset(pesticide == "atrazine") 

hu_atrazine <- 0.75 # probability of zero

dist_from_LOQ_atrazine <- 1.4

mu_atrazine <- atrazine_detect$LOQ * dist_from_LOQ_atrazine # fake atrazine mean, located 40% further from LOQ

sigma_atrazine <- 2

left_censor_atrazine <- atrazine_detect$LOD

interval_censor_atrazine <- atrazine_detect$LOQ

a_atrazine <- calculate_a_lnorm(mean = mu_atrazine, sd = sigma_atrazine)

b_atrazine <- calculate_b_lnorm(mean = mu_atrazine, sd = sigma_atrazine)


data_atrazine <- tibble(
  y = rlnorm(n1, a_atrazine, b_atrazine),
  hurdle_only = (1 - rbinom(n1, 1, prob = hu_atrazine)) * rlnorm(n1, a_atrazine, b_atrazine)
)

y_atrazine <- ggplot(data_atrazine) +
  geom_histogram(aes(x = y), binwidth = 0.1) +
  geom_vline(xintercept = 0, color = "red")+
  labs(title = "y")

y_atrazine_hurdle <- ggplot(data_atrazine) +
  geom_histogram(aes(x = hurdle_only), binwidth = 0.1) +
  geom_vline(xintercept = 0, color = "red")+
  labs(title = "hurdle data")

y_atrazine + y_atrazine_hurdle
```

Recover parameters
We now need a prior for the probability of zeros (hu)

set up model and simulate from prior only
```{r}
hurdle_bf_atrazine <- bf(hurdle_only  ~ 1,
                           hu ~ 1,
                           family = "hurdle_lognormal"
)

get_prior(hurdle_bf_atrazine, data = data_atrazine)

hurdle_prior_atrazine <- c(
  prior(normal(1.5,1), class = "Intercept", dpar = "hu"), # hu prior is on logit scale (for now)
  prior(normal(6,2), class = "Intercept"),
  prior(exponential(2), class = "sigma")
)

hurdle_mod_atrazine_priors <- brm(data = data_atrazine,
                      formula = hurdle_bf_atrazine,
                      prior = hurdle_prior_atrazine,
                      sample_prior = "only",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

hurdle_true_values <-  tibble(
  name = as.factor(c("b_Intercept", "b_hu_Intercept", "sigma")),
  value = c(a_atrazine, qlogis(hu_atrazine), b_atrazine)
)

hurdle_prior_draws <- as_draws_df(hurdle_mod_atrazine_priors) %>% select(b_Intercept, b_hu_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept, sigma))


ggplot() +
  stat_histinterval(data = hurdle_prior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")

```

priors for a and b not great
change priors
```{r}

hurdle_prior_atrazine <- c(
  prior(normal(1.5,1), class = "Intercept", dpar = "hu"), # hu prior is on logit scale (for now)
  prior(normal(3,4), class = "Intercept"),
  prior(exponential(0.5), class = "sigma")
)

hurdle_mod_atrazine_priors <- brm(data = data_atrazine,
                      formula = hurdle_bf_atrazine,
                      prior = hurdle_prior_atrazine,
                      sample_prior = "only",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

hurdle_prior_draws <- as_draws_df(hurdle_mod_atrazine_priors) %>% select(b_Intercept, b_hu_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept, sigma))


ggplot() +
  stat_histinterval(data = hurdle_prior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")

```


Better

Now model the data
```{r}
hurdle_mod_atrazine <- brm(data = data_atrazine,
                      formula = hurdle_bf_atrazine,
                      prior = hurdle_prior_atrazine,
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(hurdle_mod_atrazine)

broom.mixed::tidy(hurdle_mod_atrazine)

pp_check(hurdle_mod_atrazine)

hurdle_posterior_draws <- as_draws_df(hurdle_mod_atrazine) %>% select(b_Intercept, b_hu_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept, sigma))


ggplot() +
  stat_histinterval(data = hurdle_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")


```
The model is good to recover the parameters but the ppcheck function isn't really adapted to this sort of distribution

Let's compare the observed observations to the predicted ones

```{r}
hurdle_predictions <- data_atrazine %>%
  select(hurdle_only) %>%
  add_predicted_draws(hurdle_mod_atrazine, ndraws = 100)

ggplot(hurdle_predictions) +
  geom_point(aes(x = .prediction, y = hurdle_only)) +
  geom_abline()
```
The model expects more extreme values than the actual data

At least it recovers the parameters well though

Since we want to display predictions for the average, it may not be too bad that the model expects more extreme values

Let's move on to left censoring the lognormal

#2) simulate lognormal left censored data and recover parameters

Simulate left censored data (under LOD)
```{r}
data_atrazine <- data_atrazine %>%
  mutate(y_left = ifelse(y < left_censor_atrazine, left_censor_atrazine, y)) %>%
  mutate(censoring = ifelse(y < left_censor_atrazine, "left", "none")) %>% # need to specify the type of censoring for brms
  mutate(y_na = ifelse(censoring == "none", y, NA))

y_atrazine <- ggplot(data_atrazine) +
  geom_histogram(aes(x = y), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "y",
       subtitle = "original data")

y_atrazine_na <- ggplot(data_atrazine) +
  geom_histogram(aes(x = y_na), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "left censored data",
       subtitle = "the data we see")

y_atrazine_left <- ggplot(data_atrazine) +
  geom_histogram(aes(x = y_left), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "left censored data",
       subtitle = "the data we feed to brms")

y_atrazine + y_atrazine_left + y_atrazine_na
```

recover parameters
```{r}
left_bf_atrazine <- bf(y_left | cens(censoring) ~ 1,
                           family = "lognormal"
)

censoring_prior_atrazine <- c(
  prior(normal(6,2), class = "Intercept"),
  prior(exponential(2), class = "sigma")
)

left_mod_atrazine <- brm(data = data_atrazine,
                      formula = left_bf_atrazine,
                      prior = censoring_prior_atrazine,
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(left_mod_atrazine)

broom.mixed::tidy(left_mod_atrazine)

pp_check(left_mod_atrazine)


censoring_true_values <-  tibble(
  name = as.factor(c("b_Intercept", "sigma")),
  value = c(a_atrazine, b_atrazine)
)

left_posterior_draws <- as_draws_df(left_mod_atrazine) %>% select(b_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept,  sigma))


ggplot() +
  stat_histinterval(data = left_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = censoring_true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```

Perfect, let's try interval censoring now

#3) simulate lognormal interval censored data and recover parameters
```{r}
data_atrazine <- data_atrazine %>%
  mutate(y_interval = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, left_censor_atrazine, y),
         # every value between LOD and LOQ to LOD (the lower bound)
         
         upper_int_cens = ifelse( y >= left_censor_atrazine & y < interval_censor_atrazine, interval_censor_atrazine, y)) %>% 
  #specify upper bound of interval censoring
  
  mutate(censoring_interval = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, "interval", "none")) %>% 
  # need to specify the type of censoring for brms
  
  mutate(y_na_interval = ifelse(censoring_interval == "none", y, NA))


y_atrazine_na_interval <- ggplot(data_atrazine) +
  geom_histogram(aes(x = y_na_interval), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "interval censored data",
       subtitle = "the data we see")

y_atrazine_interval <- ggplot(data_atrazine) +
  geom_histogram(aes(x = y_interval), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "interval censored data",
       subtitle = "the data we feed to brms")

y_atrazine + y_atrazine_interval + y_atrazine_na_interval
```

Recover parameters
```{r}
interval_bf_atrazine <- bf(y_interval | cens(censoring_interval, upper_int_cens) ~ 1, # need to specify the upper bound of interval censoring
                           family = "lognormal"
)

interval_mod_atrazine <- brm(data = data_atrazine,
                      formula = interval_bf_atrazine,
                      prior = censoring_prior_atrazine,
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(interval_mod_atrazine)

broom.mixed::tidy(interval_mod_atrazine)

pp_check(interval_mod_atrazine)


interval_posterior_draws <- as_draws_df(interval_mod_atrazine) %>% select(b_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept,  sigma))


ggplot() +
  stat_histinterval(data = interval_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = censoring_true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```
Perfect, let's combine both types of censoring

#4) simulate lognormal left + interval censored data and recover parameters
Simulate data
```{r}
data_atrazine <- data_atrazine %>%
  mutate(y_full_cens = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, left_censor_atrazine, 
                              # every value between LOD and LOQ to LOD (the lower bound)
                        ifelse(y < left_censor_atrazine, left_censor_atrazine, y))) %>%
  # every value under LOD to LOD
  
  mutate(censoring_full = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, "interval", 
                            ifelse(y < left_censor_atrazine, "left", "none"))) %>% 
  # specify type of censoring
  
  mutate(y_na_full = ifelse(censoring_full == "none", y, NA))


y_atrazine_na_full <- ggplot(data_atrazine) +
  geom_histogram(aes(x = y_na_full), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "left + interval censored data",
       subtitle = "the data we see")

y_atrazine_full<- ggplot(data_atrazine) +
  geom_histogram(aes(x = y_full_cens), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "left + interval censored data",
       subtitle = "the data we feed to brms")

y_atrazine + y_atrazine_full + y_atrazine_na_full
```

Recover parameters
```{r}
full_cens_bf_atrazine <- bf(y_full_cens | cens(censoring_full, upper_int_cens) ~ 1, # need to specify the upper bound of interval censoring
                           family = "lognormal"
)

full_cens_mod_atrazine <- brm(data = data_atrazine,
                      formula = full_cens_bf_atrazine,
                      prior = censoring_prior_atrazine,
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(full_cens_mod_atrazine)

broom.mixed::tidy(full_cens_mod_atrazine)

pp_check(full_cens_mod_atrazine)


full_cens_posterior_draws <- as_draws_df(full_cens_mod_atrazine) %>% select(b_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept,  sigma))


ggplot() +
  stat_histinterval(data = full_cens_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = censoring_true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```
Perfect, let's add some hurdling to left censored data

#5) simulate hurdle lognormal left censored data and recover parameters
```{r}
data_atrazine <- data_atrazine %>%
  mutate(y_left_hurdle = ifelse(hurdle_only < left_censor_atrazine, left_censor_atrazine, hurdle_only)) %>%
  mutate(cens_left_hurdle = ifelse(hurdle_only < left_censor_atrazine, "left", "none")) %>% # need to specify the type of censoring for brms
  mutate(y_na_left_hurdle = ifelse(cens_left_hurdle == "none", hurdle_only, NA))


y_atrazine_na_left_hurdle <- ggplot(data_atrazine) +
  geom_histogram(aes(x = y_na_left_hurdle), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "left + hurdle censored data",
       subtitle = "the data we see")

y_atrazine_left_hurdle<- ggplot(data_atrazine) +
  geom_histogram(aes(x = y_left_hurdle), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "left + hurdle censored data",
       subtitle = "the data we feed to brms")

y_atrazine_hurdle <- ggplot(data_atrazine) +
  geom_histogram(aes(x = hurdle_only), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "hurdle data",
       subtitle = "original hurdle data")

y_atrazine_hurdle + y_atrazine_left_hurdle + y_atrazine_na_left_hurdle
```

Recover parameters
```{r}
hurdle_left_bf_atrazine <- bf(y_left_hurdle | cens(cens_left_hurdle) ~ 1, 
                              hu ~ 1,
                              family = "hurdle_lognormal"
)

hurdle_left_mod_atrazine <- brm(data = data_atrazine,
                      formula = hurdle_left_bf_atrazine,
                      prior = hurdle_prior_atrazine, # the priors here have changed to include prior for hu
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(hurdle_left_mod_atrazine)

broom.mixed::tidy(hurdle_left_mod_atrazine)

pp_check(hurdle_left_mod_atrazine)


left_hurdle_posterior_draws <- as_draws_df(hurdle_left_mod_atrazine) %>% select(b_Intercept, b_hu_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept,  sigma))


ggplot() +
  stat_histinterval(data = left_hurdle_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```


Perfect, let's try interval censored hurdle data

#6) simulate hurdle lognormal interval censored data and recover parameters
Simulate data
```{r}
data_atrazine <- data_atrazine %>%
  mutate(y_interval_hurdle = ifelse(hurdle_only >= left_censor_atrazine & hurdle_only < interval_censor_atrazine, left_censor_atrazine, hurdle_only),
         # every value between LOD and LOQ to LOD (the lower bound)
         
         upper_int_cens_hurdle = ifelse( hurdle_only >= left_censor_atrazine & hurdle_only < interval_censor_atrazine, interval_censor_atrazine, hurdle_only)) %>% 
  #specify upper bound of interval censoring
  
  mutate(censoring_interval_hurdle = ifelse(hurdle_only >= left_censor_atrazine & hurdle_only < interval_censor_atrazine, "interval", "none")) %>% 
  # need to specify the type of censoring for brms
  
  mutate(y_na_interval_hurdle = ifelse(censoring_interval_hurdle == "none", hurdle_only, NA))


y_atrazine_na_interval_hurdle <- ggplot(data_atrazine) +
  geom_histogram(aes(x = y_na_interval_hurdle), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "interval + hurdle censored data",
       subtitle = "the data we see")

y_atrazine_interval_hurdle<- ggplot(data_atrazine) +
  geom_histogram(aes(x = y_interval_hurdle), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "interval + hurdle censored data",
       subtitle = "the data we feed to brms")

y_atrazine_hurdle + y_atrazine_interval_hurdle + y_atrazine_na_interval_hurdle
```

Recover parameters
```{r}
hurdle_interval_bf_atrazine <- bf(y_interval_hurdle | cens(censoring_interval_hurdle , upper_int_cens_hurdle) ~ 1, 
                                  hu ~ 1, 
                           family = "hurdle_lognormal"
)

hurdle_interval_mod_atrazine <- brm(data = data_atrazine,
                      formula = hurdle_interval_bf_atrazine,
                      prior = hurdle_prior_atrazine, # the priors here have changed to include prior for hu
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(hurdle_interval_mod_atrazine)

broom.mixed::tidy(hurdle_interval_mod_atrazine)

pp_check(hurdle_interval_mod_atrazine)


interval_hurdle_posterior_draws <- as_draws_df(hurdle_interval_mod_atrazine) %>% select(b_Intercept, b_hu_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept,  sigma))


ggplot() +
  stat_histinterval(data = interval_hurdle_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```

Perfect, let's bring it all together

#7) bring it all together and simulate hurdle lognormal left and interval censored data and recover parameters

Simulate data
```{r}
data_atrazine <- data_atrazine %>%
  mutate(y_full_hurdle = ifelse(hurdle_only >= left_censor_atrazine & hurdle_only < interval_censor_atrazine, left_censor_atrazine, # every value between LOD and LOQ to LOD (the lower bound)
                        ifelse(hurdle_only < left_censor_atrazine, left_censor_atrazine, hurdle_only)), # every value under LOD to LOD
         
         upper_full_cens_hurdle = ifelse( hurdle_only >= left_censor_atrazine & hurdle_only < interval_censor_atrazine, interval_censor_atrazine, hurdle_only)) %>% #specify upper bound of interval censoring
  
  mutate(censoring_full_hurdle = ifelse(hurdle_only >= left_censor_atrazine & hurdle_only < interval_censor_atrazine, "interval", 
                            ifelse(hurdle_only < left_censor_atrazine, "left", "none"))) %>% # specify type of censoring
  mutate(y_na_full_hurdle = ifelse(censoring == "none", y, NA))

y_atrazine_na_full_hurdle <- ggplot(data_atrazine) +
  geom_histogram(aes(x = y_na_full_hurdle), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "left + interval + hurdle censored data",
       subtitle = "the data we see")

y_atrazine_full_hurdle<- ggplot(data_atrazine) +
  geom_histogram(aes(x = y_full_hurdle), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "left + interval + hurdle censored data",
       subtitle = "the data we feed to brms")

y_atrazine_hurdle + y_atrazine_full_hurdle + y_atrazine_na_full_hurdle
```

Recover parameters
```{r}
hurdle_full_bf_atrazine <- bf(y_full_hurdle | cens(censoring_full_hurdle , upper_full_cens_hurdle) ~ 1, 
                                  hu ~ 1, 
                           family = "hurdle_lognormal"
)

hurdle_full_mod_atrazine <- brm(data = data_atrazine,
                      formula = hurdle_full_bf_atrazine,
                      prior = hurdle_prior_atrazine, # the priors here have changed to include prior for hu
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(hurdle_full_mod_atrazine)

broom.mixed::tidy(hurdle_full_mod_atrazine)

print(hurdle_full_mod_atrazine)

pp_check(hurdle_full_mod_atrazine)


full_hurdle_posterior_draws <- as_draws_df(hurdle_full_mod_atrazine) %>% select(b_Intercept, b_hu_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept,  sigma))


ggplot() +
  stat_histinterval(data = full_hurdle_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```

Model overestimates prob of hu and parameter a

Let's see what that looks like on the probability and response scales 

```{r}
full_epred <- data_atrazine %>%
  select(y_full_hurdle) %>%
  add_epred_draws(hurdle_full_mod_atrazine, dpar = TRUE, ndraws = 1000) %>%
  ungroup() %>%
  select(.epred, hu) %>%
  pivot_longer(everything())

true_values <- tibble(
  name = c(".epred", "hu"),
  value = c(mu_atrazine, hu_atrazine)
)

ggplot() +
  stat_histinterval(data = full_epred , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
  
  

```



Let's see if we use sd from the actual data

```{r}
pest_conc <- read_excel("master_pesticides.xlsx")

pest_conc_long <- pest_conc %>% pivot_longer(cols = one.naphthol:fludioxonil, names_to = "pesticide", values_to = "recorded_concentration")

pest_master_long <- right_join(pest_conc_long, pest_detect, by = "pesticide")

pest_master_long <- pest_master_long %>%
  mutate(concentration = as.numeric(ifelse(recorded_concentration == "nd", 0, 
                                            ifelse(recorded_concentration == "<LOQ", LOD, recorded_concentration))), # NAs introduced by coercion -> from samples in 2013 where recorded_concentration == NA just remove for now
         status = ifelse(recorded_concentration == "nd", "nd", 
                                            ifelse(recorded_concentration == "<LOQ", "<LOQ", "quantified"))) %>%
  drop_na(concentration)



pest_master_long %>%
  subset(status == "quantified") %>%
  group_by(pesticide) %>%
    summarise(sd = sd(concentration),
              mean = mean(concentration),
              n = length(concentration))
```


sd for data that was quantified is 27 which is VERY different from the 2 that we used

I think the model has trouble placing data in the zero and concentration parts of the model as most of the mass of the distribution lies in the censored part of the distribution

To continue in this modeling avenue, we would need to simulate different distances between LOQ and the mean (as we already knew)

Okay...

Switch it up

# Alternative modeling method
What if we try a lognormal truncated at LOD but left censored from LOQ and a separate logistic model for the probability of detection

Workplan:
1) simulate and recover parameters for lognormal truncated at LOQ
2) simulate and recover parameters for lognormal truncated at LOD and left censored from LOQ

#1) simulate and recover parameters for lognormal truncated at LOQ
Simulate data
need to subset data from the hurdle process as we hypothesise that this is the data generating process
```{r}
truncated_data_LOQ <- data_atrazine %>%
  select(hurdle_only) %>%
  subset(hurdle_only >= interval_censor_atrazine) %>%
  mutate(lower_bound = interval_censor_atrazine)

y_atrazine_hurdle_truncated <- ggplot(truncated_data_LOQ) +
  geom_histogram(aes(x = hurdle_only), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "truncated data at LOQ",
       subtitle = "the data we feed to brms")

y_atrazine_hurdle + y_atrazine_hurdle_truncated
```


Recover parameters
```{r}
truncated_bf_atrazine <- bf(hurdle_only| trunc(lb = lower_bound) ~ 1,
                           family = "lognormal"
)


truncated_mod_atrazine <- brm(data = truncated_data_LOQ,
                      formula = truncated_bf_atrazine,
                      prior = censoring_prior_atrazine, # priors shouldn't change
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(truncated_mod_atrazine)

broom.mixed::tidy(truncated_mod_atrazine)

pp_check(truncated_mod_atrazine)

truncated_posterior_draws <- as_draws_df(truncated_mod_atrazine) %>% select(b_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept,  sigma))


ggplot() +
  stat_histinterval(data = truncated_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = censoring_true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```


Model overestimates parameter a and underestimates parameter b a bit

Let's try with left censored data

#2) simulate and recover parameters for lognormal truncated at LOD and left censored from LOQ

Simulate data
Need to subset hurdle data with LOD

```{r}
truncated_data_LOD <- data_atrazine %>%
  select(hurdle_only) %>%
  subset(hurdle_only >= left_censor_atrazine) %>%
  mutate(lower_bound = left_censor_atrazine) %>%
  mutate(y_left = ifelse(hurdle_only < interval_censor_atrazine, interval_censor_atrazine, hurdle_only)) %>%
  mutate(censoring = ifelse(hurdle_only < interval_censor_atrazine, "left", "none")) %>% # need to specify the type of censoring for brms
  mutate(y_na = ifelse(censoring == "none", hurdle_only, NA))
  

y_atrazine_censored_truncated <- ggplot(truncated_data_LOD) +
  geom_histogram(aes(x = hurdle_only), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "truncated data at LOD")

y_atrazine_censored_truncated_brms <- ggplot(truncated_data_LOD) +
  geom_histogram(aes(x = y_left), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "truncated data at LOD",
       subtitle = "the data we feed brms")

y_atrazine_censored_truncated_na <- ggplot(truncated_data_LOD) +
  geom_histogram(aes(x = y_na), binwidth = 0.1) +
  coord_cartesian(xlim = c(0,5)) +
  labs(title = "truncated data at LOD",
       subtitle = "the data we see")

y_atrazine_hurdle +  y_atrazine_censored_truncated_brms + y_atrazine_censored_truncated_na
```

Recover parameters
```{r}
trunc_cens_bf_atrazine <- bf(y_left| trunc(lb = lower_bound) + cens(censoring) ~ 1,
                           family = "lognormal"
)


trunc_cens_mod_atrazine <- brm(data = truncated_data_LOD,
                      formula = trunc_cens_bf_atrazine,
                      prior = censoring_prior_atrazine, # priors shouldn't change
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

print(trunc_cens_mod_atrazine)

plot(truncated_mod_atrazine)

broom.mixed::tidy(truncated_mod_atrazine)

pp_check(truncated_mod_atrazine)

truncated_posterior_draws <- as_draws_df(truncated_mod_atrazine) %>% select(b_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept,  sigma))


ggplot() +
  stat_histinterval(data = truncated_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = censoring_true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```

Okay... let's put this alternative on the back burner and try setting the mean of the distribution MUCH further from LOQ. Pesticides that were quantified had a mean ranging from 1.4 to 100 times further from LOQ



#Return to censored hurdle lognormal

define parameters
```{r}
set.seed(333)

n1 <- 2000 # sample size

atrazine_detect <- pest_detect %>%
  subset(pesticide == "atrazine") 

hu_atrazine <- 0.75 # probability of zero

dist_from_LOQ_atrazine2 <- 5

mu_atrazine2 <- atrazine_detect$LOQ * dist_from_LOQ_atrazine2 

sigma_atrazine <- 2

left_censor_atrazine <- atrazine_detect$LOD

interval_censor_atrazine <- atrazine_detect$LOQ

a_atrazine2 <- calculate_a_lnorm(mean = mu_atrazine2, sd = sigma_atrazine) # need to change name

b_atrazine2 <- calculate_b_lnorm(mean = mu_atrazine2, sd = sigma_atrazine)# need to change name
```


Simulate data
```{r}
atrazine_sim <- tibble(
  y = (1 - rbinom(n1, 1, prob = hu_atrazine)) * rlnorm(n1, a_atrazine2, b_atrazine2)
) %>%
  mutate(y_cen = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, left_censor_atrazine, # every value between LOD and LOQ to LOD (the lower bound)
                        ifelse(y < left_censor_atrazine, left_censor_atrazine, y)), # every value under LOD to LOD
         
         upper_int_cens = ifelse( y >= left_censor_atrazine & y < interval_censor_atrazine, interval_censor_atrazine, y)) %>% #specify upper bound of interval censoring
  
  mutate(censoring = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, "interval", 
                            ifelse(y < left_censor_atrazine, "left", "none"))) %>% # specify type of censoring
  mutate(y_na = ifelse(censoring == "none", y, NA))


ggplot(atrazine_sim) +
  geom_histogram(aes(x = y), binwidth = 0.1) +
  labs(title = "Censored data")

```

Recover parameters
```{r}
hurdle_prior_atrazine <- c(
  prior(normal(1.5,1), class = "Intercept", dpar = "hu"), # hu prior is on logit scale (for now)
  prior(normal(3,4), class = "Intercept"),
  prior(exponential(0.5), class = "sigma")
)


hurdle_full_bf_atrazine2 <- bf(y_cen | cens(censoring , upper_int_cens) ~ 1, 
                                  hu ~ 1, 
                           family = "hurdle_lognormal"
)

hurdle_full_mod_atrazine2 <- brm(data = atrazine_sim,
                      formula = hurdle_full_bf_atrazine2,
                      prior = hurdle_prior_atrazine, 
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(hurdle_full_mod_atrazine2)

broom.mixed::tidy(hurdle_full_mod_atrazine2)

print(hurdle_full_mod_atrazine2)

pp_check(hurdle_full_mod_atrazine2)


hurdle_true_values2 <-  tibble(
  name = as.factor(c("b_Intercept", "b_hu_Intercept", "sigma")),
  value = c(a_atrazine2, qlogis(hu_atrazine), b_atrazine2) 
)

full_hurdle_posterior_draws2 <- as_draws_df(hurdle_full_mod_atrazine2) %>% select(b_Intercept, b_hu_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept,  sigma))


ggplot() +
  stat_histinterval(data = full_hurdle_posterior_draws2 , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values2, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```


TAKE THAT!!

Now, let's increase the probability of getting a zero to 90%

```{r}
hu_atrazine2 <- 0.90


atrazine_sim2 <- tibble(
  y = (1 - rbinom(n1, 1, prob = hu_atrazine2)) * rlnorm(n1, a_atrazine2, b_atrazine2)
) %>%
  mutate(y_cen = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, left_censor_atrazine, # every value between LOD and LOQ to LOD (the lower bound)
                        ifelse(y < left_censor_atrazine, left_censor_atrazine, y)), # every value under LOD to LOD
         
         upper_int_cens = ifelse( y >= left_censor_atrazine & y < interval_censor_atrazine, interval_censor_atrazine, y)) %>% #specify upper bound of interval censoring
  
  mutate(censoring = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, "interval", 
                            ifelse(y < left_censor_atrazine, "left", "none"))) %>% # specify type of censoring
  mutate(y_na = ifelse(censoring == "none", y, NA))


ggplot(atrazine_sim2) +
  geom_histogram(aes(x = y), binwidth = 0.1) +
  labs(title = "Censored data")
```

Recover parameters
```{r}


hurdle_full_mod_atrazine3 <- brm(data = atrazine_sim2,
                      formula = hurdle_full_bf_atrazine2,
                      prior = hurdle_prior_atrazine, 
                      sample_prior = "yes",
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(hurdle_full_mod_atrazine3)

broom.mixed::tidy(hurdle_full_mod_atrazine3)

print(hurdle_full_mod_atrazine3)

pp_check(hurdle_full_mod_atrazine3)

hurdle_true_values3 <-  tibble(
  name = as.factor(c("b_Intercept", "b_hu_Intercept", "sigma")),
  value = c(a_atrazine2, qlogis(hu_atrazine2), b_atrazine2) 
)
full_hurdle_posterior_draws3 <- as_draws_df(hurdle_full_mod_atrazine3) %>% select(b_Intercept, b_hu_Intercept, sigma) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept,  sigma))


ggplot() +
  stat_histinterval(data = full_hurdle_posterior_draws3 , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values3, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```

Would you look at that!

Now let's see how the model behaves with predictors

# Censored hurdle with paysage


Simulate paysage effect
First, paysage affects only the mean
```{r}
beta_paysage <- log(1.4)

atrazine_sim_paysage <- tibble(
  paysage = seq(from = -3, to = 3, length.out = n1),
  y = (1 - rbinom(n1, 1, prob = hu_atrazine2)) * rlnorm(n1, a_atrazine2 + beta_paysage*paysage, b_atrazine2)
) %>%
  mutate(y_cen = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, left_censor_atrazine, # every value between LOD and LOQ to LOD (the lower bound)
                        ifelse(y < left_censor_atrazine, left_censor_atrazine, y)), # every value under LOD to LOD
         
         upper_int_cens = ifelse( y >= left_censor_atrazine & y < interval_censor_atrazine, interval_censor_atrazine, y)) %>% #specify upper bound of interval censoring
  
  mutate(censoring = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, "interval", 
                            ifelse(y < left_censor_atrazine, "left", "none"))) %>% # specify type of censoring
  mutate(y_na = ifelse(censoring == "none", y, NA))

ggplot(atrazine_sim_paysage, aes(x = paysage, y = y_na)) +
  geom_point()
```


Recover parameters
```{r}
hurdle_paysage_bf_atrazine <- bf(y_cen | cens(censoring , upper_int_cens) ~ 1 + paysage, 
                                  hu ~ 1, 
                           family = "hurdle_lognormal"
)


get_prior(hurdle_paysage_bf_atrazine, data = atrazine_sim_paysage)

hurdle_prior_atrazine_paysage <- c(
  prior(normal(1.5,1), class = "Intercept", dpar = "hu"), # hu prior is on logit scale (for now)
  prior(normal(3,4), class = "Intercept"),
  prior(normal(0,2), class = "b"),
  prior(exponential(0.5), class = "sigma")
)

hurdle_paysage_mod_atrazine <- brm(data = atrazine_sim_paysage,
                      formula = hurdle_paysage_bf_atrazine,
                      prior = hurdle_prior_atrazine_paysage, 
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(hurdle_paysage_mod_atrazine)

print(hurdle_paysage_mod_atrazine)

pp_check(hurdle_paysage_mod_atrazine)


hurdle_true_values <-  tibble(
  name = as.factor(c("b_Intercept", "b_hu_Intercept", "sigma", "b_paysage")),
  value = c(a_atrazine2, qlogis(hu_atrazine2), b_atrazine2, beta_paysage) 
)

full_paysage_posterior_draws <- as_draws_df(hurdle_paysage_mod_atrazine) %>% select(b_Intercept, b_hu_Intercept, sigma, b_paysage) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept,  sigma, b_paysage))


ggplot() +
  stat_histinterval(data = full_paysage_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```

This works quite well!

Let's add paysage as a predictor for the zero-probability process

In this simulation, I'll make beta_paysage for the mean negative and beta_paysage_hu positive such that as paysage increases, the mean decreases and "presence" decreases 

```{r}
beta_paysage2 <- log(0.8)

beta_paysage_hu2 <- 0.1

atrazine_sim_paysage2 <- tibble(
  paysage = seq(from = -3, to = 3, length.out = n1),
  y = (1 - rbinom(n1, 1, prob = plogis(hu_atrazine2 + beta_paysage_hu2*paysage))) * 
    rlnorm(n1, a_atrazine2 + beta_paysage2*paysage, b_atrazine2)
) %>%
  mutate(y_cen = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, left_censor_atrazine, 
                        # every value between LOD and LOQ to LOD (the lower bound)
                        ifelse(y < left_censor_atrazine, left_censor_atrazine, y)), 
         # every value under LOD to LOD
         
         upper_int_cens = ifelse( y >= left_censor_atrazine & y < interval_censor_atrazine, interval_censor_atrazine, y)) %>% 
  #specify upper bound of interval censoring
  
  mutate(censoring = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, "interval", 
                            ifelse(y < left_censor_atrazine, "left", "none"))) %>% 
  # specify type of censoring
  mutate(y_na = ifelse(censoring == "none", y, NA))

ggplot(atrazine_sim_paysage2, aes(x = paysage, y = y_na)) +
  geom_point()

ggplot(atrazine_sim_paysage2, aes(x = paysage, y = y)) +
  geom_point()
```


Recover parameters
```{r}
hurdle_paysage_bf_atrazine2 <- bf(y_cen | cens(censoring , upper_int_cens) ~ 1 + paysage, 
                                  hu ~ 1 + paysage, 
                           family = "hurdle_lognormal"
)


get_prior(hurdle_paysage_bf_atrazine2, data = atrazine_sim_paysage2)

hurdle_prior_atrazine_paysage2 <- c(
  prior(normal(1.5,1), class = "Intercept", dpar = "hu"), # hu prior is on logit scale (for now)
  prior(normal(0,2), class = "b", dpar = "hu"),
  prior(normal(3,4), class = "Intercept"),
  prior(normal(0,2), class = "b"),
  prior(exponential(0.5), class = "sigma")
)

hurdle_paysage_mod_atrazine2 <- brm(data = atrazine_sim_paysage2,
                      formula = hurdle_paysage_bf_atrazine2,
                      prior = hurdle_prior_atrazine_paysage2, 
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(hurdle_paysage_mod_atrazine2)

print(hurdle_paysage_mod_atrazine2)

pp_check(hurdle_paysage_mod_atrazine2)


full_paysage_posterior_draws2 <- as_draws_df(hurdle_paysage_mod_atrazine2) %>% select(b_Intercept, b_hu_Intercept, sigma, b_paysage, b_hu_paysage) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept, sigma, b_paysage, b_hu_paysage))

hurdle_true_values2 <-  tibble(
  name = as.factor(c("b_Intercept", "b_hu_Intercept", "sigma", "b_paysage", "b_hu_paysage")),
  value = c(a_atrazine2, hu_atrazine2, b_atrazine2, beta_paysage2, beta_paysage_hu2) 
)

ggplot() +
  stat_histinterval(data = full_paysage_posterior_draws2 , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values2, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```

The model performs really well

Let's add % cover water

# (PINNED) Censored hurdle with paysage and % cover water

water affects only the mean
```{r}

beta_paysage2 <- log(0.8)

beta_paysage_hu2 <- 0.1

beta_water <- log(1.4)

atrazine_sim_water <- tibble(
  paysage = seq(from = -3, to = 3, length.out = n1),
  water = seq(from = 0, to = 1,  length.out = n1),
  y = (1 - rbinom(n1, 1, prob = plogis(hu_atrazine2 + beta_paysage_hu2*paysage))) * 
    rlnorm(n1, a_atrazine2 + beta_paysage2*paysage + beta_water*water, b_atrazine2)
) %>%
  mutate(y_cen = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, left_censor_atrazine, 
                        # every value between LOD and LOQ to LOD (the lower bound)
                        ifelse(y < left_censor_atrazine, left_censor_atrazine, y)), 
         # every value under LOD to LOD
         
         upper_int_cens = ifelse( y >= left_censor_atrazine & y < interval_censor_atrazine, interval_censor_atrazine, y)) %>% 
  #specify upper bound of interval censoring
  
  mutate(censoring = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, "interval", 
                            ifelse(y < left_censor_atrazine, "left", "none"))) %>% 
  # specify type of censoring
  mutate(y_na = ifelse(censoring == "none", y, NA))

ggplot(atrazine_sim_water, aes(x = paysage, y = y_na)) +
  geom_point()

ggplot(atrazine_sim_water, aes(x = water, y = y_na)) +
  geom_point()

ggplot(atrazine_sim_water, aes(x = paysage, y = y)) +
  geom_point()

ggplot(atrazine_sim_water, aes(x = water, y = y)) +
  geom_point()
```


Recover parameters
```{r}
hurdle_water_bf_atrazine <- bf(y_cen | cens(censoring , upper_int_cens) ~ 1 + paysage + water, 
                                  hu ~ 1 + paysage, 
                           family = "hurdle_lognormal"
)


get_prior(hurdle_water_bf_atrazine, data = atrazine_sim_water)

hurdle_prior_atrazine_water <- c(
  prior(normal(1.5,1), class = "Intercept", dpar = "hu"), # hu prior is on logit scale (for now)
  prior(normal(0,2), class = "b", dpar = "hu"),
  prior(normal(3,4), class = "Intercept"),
  prior(normal(0,2), class = "b"),
  prior(exponential(0.5), class = "sigma")
)

hurdle_paysage_mod_water <- brm(data = atrazine_sim_water,
                      formula = hurdle_water_bf_atrazine,
                      prior = hurdle_prior_atrazine_water, 
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(hurdle_paysage_mod_water)

print(hurdle_paysage_mod_water)



water_posterior_draws <- as_draws_df(hurdle_paysage_mod_water) %>% select(b_Intercept, b_hu_Intercept, sigma, b_paysage, b_hu_paysage, b_water) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept, sigma, b_paysage, b_hu_paysage, b_water))

hurdle_true_values3 <-  tibble(
  name = as.factor(c("b_Intercept", "b_hu_Intercept", "sigma", "b_paysage", "b_hu_paysage", "b_water")),
  value = c(a_atrazine2, hu_atrazine2, b_atrazine2, beta_paysage2, beta_paysage_hu2, beta_water) 
)

ggplot() +
  stat_histinterval(data = water_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values3, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```

Hmm...

I think the model is unsure of parameter values due to the predictors not being scaled

# (PINNED) Explore simulating scaled parameters

Let's just see how the model behaves when I scale the predictors for the mean

```{r}
scaled_water_bf_atrazine <- bf(y_cen | cens(censoring , upper_int_cens) ~ 1 + scale(paysage) + scale(water), 
                                  hu ~ 1 + paysage, 
                           family = "hurdle_lognormal"
)


get_prior(scaled_water_bf_atrazine, data = atrazine_sim_water)

hurdle_prior_atrazine_water <- c(
  prior(normal(1.5,1), class = "Intercept", dpar = "hu"), # hu prior is on logit scale (for now)
  prior(normal(0,2), class = "b", dpar = "hu"),
  prior(normal(3,4), class = "Intercept"),
  prior(normal(0,2), class = "b"),
  prior(exponential(0.5), class = "sigma")
)

scaled_paysage_mod_water <- brm(data = atrazine_sim_water,
                      formula = scaled_water_bf_atrazine,
                      prior = hurdle_prior_atrazine_water, 
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(scaled_paysage_mod_water)

print(scaled_paysage_mod_water)



scaled_water_posterior_draws <- as_draws_df(scaled_paysage_mod_water) %>% select(b_Intercept, b_hu_Intercept, sigma, b_scalepaysage, b_hu_paysage, b_scalewater) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept, sigma, b_scalepaysage, b_hu_paysage, b_scalewater))

hurdle_true_values3 <-  tibble(
  name = as.factor(c("b_Intercept", "b_hu_Intercept", "sigma", "b_scalepaysage", "b_hu_paysage", "b_scalewater")),
  value = c(a_atrazine2, hu_atrazine2, b_atrazine2, beta_paysage2, beta_paysage_hu2, beta_water) 
)

ggplot() +
  stat_histinterval(data = scaled_water_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values3, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```


Let's put a pin in that and try simulating different pesticides with random intercepts and then slopes

Use the LOD and LOQ values for the other pesticides
# Random effects

##Random intercepts

Paysage affects just the mean
```{r}
n_per_pesticide <- 2000

pesticides <- pest_detect %>% mutate(pesticide_num = as.numeric(as.factor(pesticide))) %>% select(pesticide_num)

pesticide_intercepts <- rnorm(n = length(pesticides$pesticide_num), mean = 0, sd = log(1.2)) # pesticides vary by 40% below and above the mean (2*sd)

beta_paysage2 <- log(0.8)



random_intercepts <- tibble(
  pesticide_ID = pesticides$pesticide_num,
  pesticide_offset = pesticide_intercepts[pesticides$pesticide_num]
) %>% 
  
  expand_grid(paysage = seq(from = -3, to = 3, length.out = n_per_pesticide)) %>%
  
  mutate(y = (1 - rbinom(length(paysage), 1, prob = plogis(hu_atrazine2))) * 
    rlnorm(length(paysage), a_atrazine2 + beta_paysage2*paysage + pesticide_offset, b_atrazine2)
) %>%
  
  mutate(y_cen = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, left_censor_atrazine, 
                        # every value between LOD and LOQ to LOD (the lower bound)
                        ifelse(y < left_censor_atrazine, left_censor_atrazine, y)), 
         # every value under LOD to LOD
         
         upper_int_cens = ifelse( y >= left_censor_atrazine & y < interval_censor_atrazine, interval_censor_atrazine, y)) %>% 
  #specify upper bound of interval censoring
  
  mutate(censoring = ifelse(y >= left_censor_atrazine & y < interval_censor_atrazine, "interval", 
                            ifelse(y < left_censor_atrazine, "left", "none"))) %>% 
  # specify type of censoring
  mutate(y_na = ifelse(censoring == "none", y, NA))


ggplot(random_intercepts) +
  geom_smooth(aes(x = paysage, y = y, color = as.factor(pesticide_ID)), method = "lm", se = FALSE) +
  geom_smooth(aes(x = paysage, y = y), color = "black", method = "lm", se = FALSE, size = 2)

```


Recover parameters

```{r}
random_intercept_bf <- bf(y_cen | cens(censoring , upper_int_cens) ~ 1 + paysage + (1|pesticide_ID), 
                                  hu ~ 1 + paysage, 
                           family = "hurdle_lognormal"
)


get_prior(random_intercept_bf, data = random_intercepts)

hurdle_prior_atrazine_water <- c(
  prior(normal(1.5,1), class = "Intercept", dpar = "hu"), # hu prior is on logit scale (for now)
  prior(normal(0,2), class = "b", dpar = "hu"),
  prior(normal(3,4), class = "Intercept"),
  prior(normal(0,2), class = "b"),
  prior(exponential(0.5), class = "sigma")
)

random_intercept_mod <- brm(data = random_intercepts,
                      formula = random_intercept_bf,
                      prior = hurdle_prior_atrazine_water, 
                      iter = 2000, warmup = 1000, chains = 4, cores = 4,
                      seed = 12,
                      backend = "cmdstanr")

plot(random_intercept_mod)

print(random_intercept_mod)



scaled_water_posterior_draws <- as_draws_df(random_intercept_mod) %>% select(b_Intercept, b_hu_Intercept, sigma, b_scalepaysage, b_hu_paysage, b_scalewater) %>%
  pivot_longer(c(b_Intercept, b_hu_Intercept, sigma, b_scalepaysage, b_hu_paysage, b_scalewater))

hurdle_true_values3 <-  tibble(
  name = as.factor(c("b_Intercept", "b_hu_Intercept", "sigma", "b_scalepaysage", "b_hu_paysage", "b_scalewater")),
  value = c(a_atrazine2, hu_atrazine2, b_atrazine2, beta_paysage2, beta_paysage_hu2, beta_water) 
)

ggplot() +
  stat_histinterval(data = scaled_water_posterior_draws , aes(x = value, y = 0), point_interval = mode_hdi, .width = .95) + 
  geom_vline(data = hurdle_true_values3, aes(xintercept = value), color = "red") +
  facet_wrap(~name, scales = "free")
```





